Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2018-07-16 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates

%Viscous flow instabilities can be found across several disciplines and scales, from chemical separation processes to geological reservoir fluids. 
\citet{muskat_1934} used Hele-Shaw cells (\ie parallel flat plates separated by an infinitesimal gap) to study fluid flow and the impact on the capillary number $\left(\text{N}_{c}\right)$ in the flow dynamics.
%investigated fluid flow in Hele-Shaw cells, \ie parallel flat plates separated by an infinitesimal gap, and the impact on the capillary number ($N_{c}$),
%\begin{equation}
%N_{c} = \frac{\mu U}{\gamma}.\label{eq:capillary_number}
%\end{equation}
 %In Eqn.~\ref{eq:capillary_number}, $\gamma$ is the surface tension and $U$ is the characteristic velocity of the moving interface. 
This experimental apparatus enabled instabilities to be qualitatively investigated by simplifying the flow (in both porous and non-porous media) into a 2D problem. \citet{mclean_1981} developed a semi-analytic solution for flows in Hele-Shaw cells which were later used by \citet{guan_2003} to investigate fingers' formation, dimensions and branchiness.

More recently, \citet{howison_2000} and \citet{praud_2005} provided a comprehensive description of the mathematical formulation of immiscible two-phase flows in Hele-Shaw cells. For a Hele-Shaw cell of a given size, flow development depends only on the capillary number, therefore if \textit{$N_{c}$} is too high, \citet{saffman_1959b} \citep[see also][]{saffman_1959,homsy_1987,tabeling_1987} demonstrated that the flow develops to a single steady-state finger which moves throughout the cell with constant velocity.

>>>>

%Viscous flow instabilities can be found across several disciplines and scales, from chemical separation processes to geological reservoir fluids. 
\citet{muskat_1934} used Hele-Shaw cells (\ie parallel flat plates separated by an infinitesimal gap) to study fluid flow and the impact on the capillary number $\left(\text{N}_{c}\right)$ in the flow dynamics.
%investigated fluid flow in Hele-Shaw cells, \ie parallel flat plates separated by an infinitesimal gap, and the impact on the capillary number ($N_{c}$),
%\begin{equation}
%N_{c} = \frac{\mu U}{\gamma}.\label{eq:capillary_number}
%\end{equation}
 %In Eqn.~\ref{eq:capillary_number}, $\gamma$ is the surface tension and $U$ is the characteristic velocity of the moving interface. 
This experimental apparatus enabled instabilities to be qualitatively investigated by simplifying the flow (in both porous and non-porous media) into a 2D problem. \citet{mclean_1981} developed a semi-analytic solution for flows in Hele-Shaw cells which was later used by \citet{guan_2003} to investigate fingers' formation, dimensions and branchiness.

More recently, \citet{howison_2000} and \citet{praud_2005} provided a comprehensive description of the mathematical formulation of immiscible two-phase flows in Hele-Shaw cells. For a Hele-Shaw cell of a given size, flow development depends only on the capillary number, therefore if \textit{$N_{c}$} is too high, \citet{saffman_1959b} \citep[see also][]{saffman_1959,homsy_1987,tabeling_1987} demonstrated that the flow develops to a single steady-state finger which moves throughout the cell with constant velocity. The impact of viscosity ratio and mass flow-rate on fingering development was experimentally (coreflood and micromodels) studied by \citet{Doorwar_2016}. Through numerical (based on pore-network model) experiments, they demonstrated that fingers' dynamics (\ie growth, splitting and coalescence) are also strongly dependent on the domain's aspect-ratio and capillary forces \citep[see][]{Doorwar_2014}, and proposed a new instability dimensionless number to assess breakthrough recovery of a number of experimental data sets.
